---
title: "Tidy Tuesday 3"
author: "Shelbie Ishimaru"
date: "2024-11-09"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message= FALSE, #No messages as the default for all code chunks
                      warning= FALSE, #No warnings as the default for all code chunks
                      fig.path= "../Output/") #creates file path to save all plots
```

#### What I learned this week!
This week I learned about a fun color palette called [feathers](https://github.com/shandiya/feathers) that is based off of female Australian bird plumage. I also learned how to calculate Pearson's Correlation Coefficient in R and add the result to a scatter plot. Finally I used the patchwork package that we learned about in week 8 to combined two plots.

#### Load Libraries
```{r}
library(rnaturalearth) #for creating maps
library(rnaturalearthdata) #for creating maps 
library(tidyverse) #for data manipulation
library(countries) #for creating maps
library(patchwork) #for joining plots
library(feathers) #a fun color palette based on female Australian bird plumage
```

#### Read-in Data
```{r}
countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/countries.csv') #read in country data

country_subdivisions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/country_subdivisions.csv') #read in country subdivision data

world <- ne_countries(scale = "medium", returnclass = "sf") #read in world data 
#head(world) #look at data, commented out because its a large df
```

#### Data Manipulation
```{r}
world <- world %>% #pull in world data
  filter(pop_year== 2019) %>% #subset data to only include 2019 instances
  mutate(economy= str_replace(economy, "1. Developed region: G7", "Developed region: G7")) %>% #edit wording for final plot
  mutate(economy= str_replace(economy, "2. Developed region: nonG7", "Developed region: nonG7")) %>% #edit wording for final plot
  mutate(economy= str_replace(economy, "3. Emerging region: BRIC", "Emerging region: BRIC")) %>% #edit wording for final plot
  mutate(economy= str_replace(economy, "4. Emerging region: MIKT", "Emerging region: MIKT")) %>% #edit wording for final plot
  mutate(economy= str_replace(economy, "5. Emerging region: G20", "Emerging region: G20")) %>% #edit wording for final plot
  mutate(economy= str_replace(economy, "6. Developing region", "Developing region")) %>% #edit wording for final plot
  mutate(economy= str_replace(economy, "7. Least developed region", "Least developed region")) %>% #edit wording for final plot
  mutate(economy= factor(economy, levels= c("Developed region: G7", "Developed region: nonG7", "Emerging region: BRIC", "Emerging region: MIKT", "Emerging region: G20", "Developing region", "Least developed region"))) #edit wording for final plot

country_data <- country_subdivisions %>% #pull in country subdivision data
  group_by(alpha_2) %>% #group instances by country
  summarise(total_subdivs= n()) %>% #calculate each countries total subdivisions
  left_join(countries) %>% #join subdivision with country data
  rename(admin= name) %>% #rename the column for 2nd join
  inner_join(world) %>% #join country data with world data
  select(admin, total_subdivs, pop_est, economy) #select only the columns we want for analysis
```

#### Data Analysis
```{r}
r <- cor.test(country_data$total_subdivs, country_data$pop_est) #run Pearson's correlation test
r #look at result
```

#### Create Plot
```{r plot1}
plot1 <- ggplot(country_data, aes(x= log(total_subdivs), y= log(pop_est), color= economy)) + #create plot 
  geom_point() + #make it a scatter plot
  labs(title= "Investigating the Relationship Between Country Subdivisions, Population Size, and Economy in 2019", #add plot title
       x= "log(Population Estimate)", #add x-axis title
       y= "log(Total Subdivisions)") + #add y-axis title
  scale_color_manual(values= get_pal("bee_eater")) + #a fun color palette based on female Australian bird plumage
  theme_bw() + #change theme so we have a white background with grey grid lines and black outline
  theme(legend.position = "none", #remove legend
      plot.title= element_text(face= "bold", size= 14), #change text size and bold title
      axis.text= element_text(size= 9), #change size of axis values
      plot.caption= element_text(size= 7)) + #change text size of caption
  annotate("text", x= 1.3, y= 24, label= paste0("r = ", round(r$estimate, 3)), color= "black") #add Pearson's r value to the top corner 
plot1
```

```{r plot2}
plot2 <- ggplot(country_data, aes(x= log(total_subdivs), y= log(pop_est), color= economy)) +  #create plot 
  geom_point() + #make it a scatter plot
  facet_wrap(~economy) + #make an individual plot by economic class
  scale_color_manual(values= get_pal("bee_eater")) + #a fun color palette based on female Australian bird plumage
  labs(title= "", #remove plot title
       x= "", #remove x-axis title
       y= "", #remove y-axis title
       caption= "Source: Data Science Learning Community. 2024. Tidy Tuesday: A weekly social data project.") + #add caption that gives data credit
  theme_bw() + #change theme so we have a white background with grey grid lines and black outline
  theme(legend.position = "none") #remove legend
plot2
```

```{r final_plot, fig.width=12, fig.height=7}
plot1 + plot2 #join the two plots together to create the final plot
```

