---
title: "Week 6 Homework"
author: "Shelbie Ishimaru"
date: "2024-10-07"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message= FALSE, #No messages as the default for all code chunks
                      warning= FALSE, #No warnings as the default for all code chunks
                      fig.path= "../Outputs/") #creates file path to save all plots 
```
In this RMarkdown file, we will run through how to...

1. Manipulate a data frame efficiently using the tidyverse 
2. Create a table using the *kableExtra* package 
3. Make a figure using ggplot  
4. Utilize fun color palettes to make **BEAUTIFUL** figures    

So let's get this party started! 


### Load Libraries  
The first thing we want to do is load our libraries. Below are all of the important libraries needed to accomplish our goals and a short description of what they do.
```{r}
library(tidyverse) #for data manipulation
library(here) #for unbreakable file paths
library(kableExtra) #for building tables 
library(beyonce) #a fun color palette
```


### Load and Inspect the Data
Next we want to load in our data. We are going to be using chemical data collected from *Silbiger et al. 2020*. Once we load the data we want to look at it so we can understand what information we are working with.
```{r}
ChemData<-read_csv(here("Week_06","Data", "chemicaldata_maunalua.csv")) #read in new data set
glimpse(ChemData) #look at the data frame
```


### Data Manipulation
Now that the data is loaded we can begin manipulating our data so that it can be used for creating tables and figures. Below I am creating a clean data frame by removing *NA* values, pivoting the data, and summarizing to obtain mean nutrient concentrations by sampling zone.
```{r}
chem_clean <- ChemData %>% #creates a new data frame and pulls in our data frame
drop_na() %>% #filters out everything that is not a complete row
  select(Zone, NN, Silicate, Phosphate) %>% #selects the cols we are interested for calculations
  pivot_longer(cols = NN:Phosphate, #the cols you want to pivot. This says select the NN, silicate, and phosphate cols
               names_to = "Variables", #the names of the new cols with all the column names
               values_to = "Values") %>% #names of the new column with all the values
  group_by(Variables, Zone) %>% #groups by our desired variables
  summarise(Mean = mean(Values, na.rm = T)) #calculates the mean NN, silicate, and phosphate by zone
```


### Create a Table with Kable
After cleaning the data, we can now create a table for publications and presentations. Below I show how the *kableExtra* package can be used to make high quality tables. 
```{r}
chem_table <- chem_clean %>% #creates a new data frame and pulls in our data frame
   kbl()  %>% # make it a kable table 
  kable_classic_2(lightable_options = "striped") %>% #makes every other row a grey background
  kable_styling(full_width = FALSE, row_label_position = "c") #make table narrower and centers the row headers
chem_table #call the table so it can be shown in the markdown HTML
```


### Make a Plot
Finally we can make a figure ggplot. Below I will show how to create a plot that shows changes in Nitrate+Nitrite, Phosphate, and Silicate by Sampling Zone. To add excitement to my plot, I used the [Beyonce color palette](https://github.com/dill/beyonce)!
```{r chem_plot}
chem_clean %>%
  ggplot(aes(x = Mean, #x-axis is mean variable concentration
             y = Zone, fill= Zone)) + #y-axis is sampling zone
  geom_bar(stat= "identity", #plot the given means by zone
           position= position_dodge(), #creates clustered bar plot (by zone) 
           show.legend = F) + #removes legend
  facet_wrap(~Variables, scales= "free_x", labeller= ) +
  labs(x= "Mean Variable Concentration (umol/L)", #x-axis label
       y= "Sampling Zone", #y-axis label
       title= "Changes in Nitrate+Nitrite (NN), Phosphate, and Silicate by Sampling Zone", #plot title 
       caption= "Source: Silbiger et al. 2020") + #adds caption of where we got the data
  scale_fill_manual(values= beyonce_palette(11)) + #changes colors of bars to be Beyonce themed!
  theme_bw() + #change theme so we have a white background with grey grid lines and black outline
    theme(axis.title= element_text(size= 11), #change axes text size
        plot.title= element_text(hjust= 0.5), #change title text size
        plot.caption.position = "plot") #make caption on the bottom right of the page
```

Thanks for reading!! If you enjoyed this Markdown file, checkout my [MBIO 612 github repository](https://github.com/OCN-682-UH/Ishimaru) to see what I've been learning! :)